### 2018-04-04 latex/icpp18/abstract.tex
made updates to abstract and intro based on comments from ramki and oguz

\begin{abstract}
Non-negative matrix factorization (NMF) is the problem of finding two non-negative low rank factors $\WW$ and $\HH$ for a given input matrix $\AA$ such that $\AA \approx \WW \HH$.
NMF is a useful tool for many internet applications such as topic modeling in text mining and community detection in social networks.
In this paper, we focus on scaling distributed NMF to very large sparse datasets by employing effective algorithms, communications, and partitioning schemes that leverage the sparsity of the input matrix $\AA$.
Our method employs an optimal point-to-point communication scheme, and removes any redundant communication existing in the state of the art.
Unlike the state of the art, our method also permits arbitrary partitions of the sparse input matrix as well as factor matrices, which enables us to employ effective partitioning schemes for better scalability.
We experiment on a cluster using up to 3072 cores, and report significant better scalability with respect to existing methods owing to our efficient communication and partitioning strategies.
We also run our algorithms on an IBM BlueGene/Q supercomputer, and achieve scalability up to 32768 processors with a 32x speedup over the execution using 256 processors.

>>>>

%!TEX root = icpp18.tex

\begin{abstract}
Non-negative matrix factorization (NMF), the problem of finding two non-negative low-rank factors whose product approximates an input matrix, is a useful tool for many data mining and scientific applications such as topic modeling in text mining and blind source separation in microscopy.
In this paper, we focus on scaling algorithms for NMF to very large sparse datasets and massively parallel machines by employing effective algorithms, communication patterns, and partitioning schemes that leverage the sparsity of the input matrix.
In particular, we consider two previous works developed for related problems, one that uses a fine-grained partitioning strategy using a point-to-point communication pattern and on that uses a coarse-grained partitioning strategy using a collective-based communication pattern.
We show that a combination of the previous approaches balances the demands of the various computations within NMF algorithms and achieves high efficiency and scalability.
\grey{To be completed...

We present experiments on (THESE DATASETS), running on up to (YY) cores of a Cray XK7 supercomputer.
(MAYBE STATE SIMPLE VERSION OF CONCLUSIONS?)
}

### 2018-04-04 latex/icpp18/introduction.tex
made updates to abstract and intro based on comments from ramki and oguz

% How NMF is computed, and in sparse what is important.
The most common method for solving \cref{eqn:origNMF} is to use an alternating optimization approach, iteratively updating $\WW$ with $\HH$ fixed and then updating $\HH$ with $\WW$ fixed.
Specifically, updating a factor matrix involves three main operations; computing of $\WW^T\AA$ or $\AA\Hm^T$, computing the Gram matrix $\Wm^T \Wm$ or $\Hm \Hm^T$, and solving a non-negative linear least squares~(NNLS) problem using these two resulting matrices to update the factor matrix.
The typical bottleneck is the computation that involves the input matrix, $\WW^T\AA$ and $\AA\Hm^T$, which are sparse-dense matrix multiplications (SpMMs) when the input data is sparse.

@@

HyperTensor solves an input-sparsity-dependent hypergraph partitioning problem to determine the parallel distribution of the entries of $\AA$.
The advantage of this approach is that by being sensitive to the nonzero pattern, the resulting factor matrix communication required to compute the SpMMs is (approximately) minimized, and the only data that a processor receives is what it needs for its local computation.
The disadvantages of this approach are an upfront partitioning cost (which can be amortized over the iterations of the NMF algorithm) and possible load imbalance, in terms of the rest of the computation and the number of messages communicated among processors.

>>>>

% How NMF is computed, and in sparse what is important.
The most common method for solving \cref{eqn:origNMF} is to use an alternating optimization approach, iteratively updating $\WW$ with $\HH$ fixed and then updating $\HH$ with $\WW$ fixed.
Specifically, updating a factor matrix involves three main operations; computing of $\WW^T\AA$ or $\AA\Hm^T$, computing the Gram matrix $\Wm^T \Wm$ or $\Hm \Hm^T$, and solving a non-negative linear least squares (NLS) problem using these two resulting matrices to update the factor matrix.
When the rank $k$ is small, the typical bottleneck is the computation that involves the input matrix, $\WW^T\AA$ and $\AA\Hm^T$, which are sparse-dense matrix multiplications (SpMMs) when the input data is sparse.
However, for large $k$, the Gram and NLS computations can become the bottleneck, as they grow more quickly with $k$ than the SpMMs.
Efficient parallel algorithms for NMF must load balance the computation and avoid communication overheads.
The distribution of the input $\AA$ across processors affects the load balance and communication of the SpMMs, and the distribution of the factor matrices $\WW$ and $\HH$ affects the load balance of the Gram and NLS computations.

@@

HyperTensor solves an input-sparsity-dependent hypergraph partitioning problem to determine the parallel distribution of the entries of $\AA$.
The advantage of this approach is that by being sensitive to the nonzero pattern, the resulting factor matrix communication required to compute the SpMMs is (approximately) minimized, and the only data that a processor receives is what it needs for its local computation.
The disadvantages of this approach are an upfront partitioning cost (which can be amortized over the iterations of the NMF algorithm) and possible load imbalance, particularly in the NLS computations.
